---
title: "727 Final Project"
author: "Lingxi Li"
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(xml2)
library(rvest)
library(robotstxt)
library(tidyverse)
library(magrittr)
```
Test if web scraping is allowed
```{r}
paths_allowed("https://www.goodreads.com/shelf/show/race?page=1")
```
This web requires login to loop through all pages
```{r}
#Address of the login webpage
login<-"https://www.goodreads.com/user/sign_in?returnurl=%2Fshelf%2Fshow%2Frace%3Fpage%3D1"

#create a web session with the desired login address
pgsession<-html_session(login)
pgform<-html_form(pgsession)[[1]] #in this case the submit is the 1st form
filled_form<-set_values(pgform,'user[email]'="***", 'user[password]'= "***")
submit_form(pgsession, filled_form)
```
####Books about race
create loop to get average rating, number of rating, published year
```{r}
PAGE<-25
f_pubyr<-data.frame()

for (i in 1:PAGE) {
  url <- paste0("https://www.goodreads.com/shelf/show/race?page=",i)
  print(url)
  page<-jump_to(pgsession, url)
  nds <- html_nodes(page, xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "greyText", " " )) and contains(concat( " ", @class, " " ), concat( " ", "smallText", " " ))]')
  pub_yr <- html_text(nds)
  pub_yr<-as.data.frame(pub_yr)
  f_pubyr<-rbind(f_pubyr,pub_yr)
}
```
clean the data to get the "publish year"
```{r}
#view the first 6 rows
head(f_pubyr)
#remove irrelevant rows (all irrelavent rows contains ")")
a<-as.vector(f_pubyr$pub_yr)
length(which(grepl(")",a)==TRUE)) #1308-58 is exactly 1250, the number of rows of clean data
f_pubyr %<>% filter(grepl(")",a)==FALSE)
#extract the year
f_pubyr %<>% 
  separate(pub_yr,c("rating","num_ra","year"),"—") %>%
  separate(year,c(NA,"year"),"published")  #all "missing pieces" from the warning message is the text misscraped that does not include "publish year"
#trim line breaker
f_pubyr$year<-trimws(f_pubyr$year,which="both")
```
Now we have a single column "year" that contains all the "publish year"
```{r}
#breifly check the value of variable "year" 
table(f_pubyr$year,exclude=NULL) 
                    #29 missing because of incomplete text online                            
                    #frequency before 1960 all less than 3

#bar chart: Number of books about race per year
#filter out NA and blank
f_pubyr %>% 
  filter(year!="" & year!=1) %>%
  group_by(year) %>%
  summarise(total=n()) %>%
  ggplot(aes(year,total,fill=total))+
  geom_col(width=1.3)+
  scale_fill_continuous(low="light blue", high="blue")+
  scale_x_discrete(breaks = seq(1960, 2020, by = 5)) +
  theme(axis.text.x = element_text(angle = 45))
```
clean the data for "rating" of book about race
```{r}
#trim line breaker
f_pubyr$rating<-trimws(f_pubyr$rating,which="both")
#extract number of rating from column "rating"
f_pubyr$rating<-substr(f_pubyr$rating,12,15)
```
Now we have a single column "rating"
```{r}
#check value
table(f_pubyr$rating,exclude=NULL)
#line chart: Average rating of books about race by year
f_pubyr %>%
  filter(year!="" & year!=1) %>%
  group_by(year) %>%
  summarise(Mean_rating=mean(as.numeric(rating))) %>%
  ggplot(aes(x=year,y=Mean_rating,group=1))+
  geom_line()+
  ylim(2.75,5)+
  scale_x_discrete(breaks = seq(1600, 2020, by = 10)) +
  theme(axis.text.x = element_text(angle = 45))
```
Run another loop to get book names
```{r}
PAGE<-25
f_rana<-data.frame()

for (i in 1:PAGE) {
  url <- paste0("https://www.goodreads.com/shelf/show/race?page=",i)
  print(url)
  page <- jump_to(pgsession, url)
  nds <- html_nodes(page, xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "bookTitle", " " ))]')
  race_name <- html_text(nds)
  race_name <- as.data.frame(race_name)
  f_rana <- rbind(f_rana,race_name)
}
```
put name and year into the same dataframe
```{r}
f_race <- cbind(f_rana, f_pubyr)
```
####Top 200 Books for Each Year
get the names of top 200 books from 1960 to 2019
```{r}
YEAR<-2019  #1960-2019
f_top<-data.frame()

for (i in 1960:YEAR) {
  url <- paste0("https://www.goodreads.com/book/popular_by_date/",i)
  print(url)
  page <- jump_to(pgsession, url)
  nds <- html_nodes(page, xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "bookTitle", " " ))]//span')
  top_name <- html_text(nds)
  top_name <- as.data.frame(top_name)
  f_top <- rbind(f_top,top_name)
}
```
find intersections between top200 and books about race by year
```{r}
#trim the line breaker
f_race$race_name<-trimws(f_race$race_name,which="both")
f_top$top_name<-trimws(f_top$top_name,which="both")
#delet whatever is in parentheses
f_race$race_name<-gsub("\\s*\\([^\\)]+\\)","",f_race$race_name)
f_top$top_name<-gsub("\\s*\\([^\\)]+\\)","",f_top$top_name)
#loop to get the number of race books among the top 200 for each year
popular_inter1 <- data.frame()
for (i in 1960:2019) {
  year_group <- which(f_race$year==i)
  intersection <- intersect(f_top$top_name,f_race$race_name[year_group])
  intersection <- as.data.frame(length(intersection))
  popular_inter1 <- rbind(popular_inter1,intersection)
}
names(popular_inter1)<-"number"
#calculate the proportion
popular_inter1$proportion <- popular_inter1$number/200
#append column of year
popular_inter1$year <- c(1960:2019)
popular_inter1$year <- as.factor(popular_inter1$year)
#plot of the proportion
ggplot(popular_inter1,aes(x=year,y=proportion,fill=proportion))+
  geom_col(width=1)+
  scale_fill_continuous(low="light blue", high="blue")+
  scale_x_discrete(breaks = seq(1960, 2020, by = 5)) +
  ylim(0,0.1)+
  theme(axis.text.x = element_text(angle = 45))
```
####Books about politics
create loop to get average rating, number of rating, published year
```{r}
PAGE<-25
f_pubyr2<-data.frame()

for (i in 1:PAGE) {
  url <- paste0("https://www.goodreads.com/shelf/show/politics?page=",i)
  print(url)
  page<-jump_to(pgsession, url)
  nds <- html_nodes(page, xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "greyText", " " )) and contains(concat( " ", @class, " " ), concat( " ", "smallText", " " ))]')
  pub_yr <- html_text(nds)
  pub_yr<-as.data.frame(pub_yr)
  f_pubyr2<-rbind(f_pubyr2,pub_yr)
}
```
clean the collected data-politics
```{r}
#view the first 6 rows
head(f_pubyr2)
#remove irrelevant rows (all irrelavent rows contains ")")
f_pubyr2 %<>% filter(grepl(")",pub_yr)==FALSE)
#extract the year
f_pubyr2 %<>% 
  separate(pub_yr,c("rating","num_ra","year"),"—") %>%
  separate(year,c(NA,"year"),"published") 
#trim line breaker
f_pubyr2$year<-trimws(f_pubyr2$year,which="both")
```
Now we have a single column "year" that contains all the "publish year"
```{r}
#breifly check the value of variable "year" 
table(f_pubyr2$year,exclude=NULL) 
                    #9 missing because of incomplete text online                            
                    #frequency before 1960 all less than 5
#bar chart: Number of books about politics per year
#filter out NA and blank
f_pubyr2 %>% 
  filter(year!="" & as.numeric(year)>=1603) %>%
  group_by(year) %>%
  summarise(total=n()) %>%
  ggplot(aes(year,total,fill=total))+
  geom_col(width=1.2)+
  scale_fill_continuous(low="light blue", high="blue")+
  scale_x_discrete(breaks = seq(1960, 2020, by = 5)) +
  theme(axis.text.x = element_text(angle = 45))
```
clean the data for "rating" of book about politics
```{r}
#trim line breaker
f_pubyr2$rating<-trimws(f_pubyr2$rating,which="both")
#extract number of rating from column "rating"
f_pubyr2$rating<-substr(f_pubyr2$rating,12,15)
```
Now we have a single column "rating"
```{r}
#check value
table(f_pubyr$rating,exclude=NULL)
#bar chart: Average rating of books about race by year
f_pubyr2 %>%
  filter(year!="" & as.numeric(year)>0) %>%
  group_by(year) %>%
  summarise(Mean_rating=mean(as.numeric(rating))) %>%
  ggplot(aes(x=year,y=Mean_rating,group=1))+
  geom_line()+
  ylim(2.75,5)+
  scale_x_discrete(breaks = seq(1600, 2020, by = 10)) +
  theme(axis.text.x = element_text(angle = 45))
```
Run another loop to get book names-politics
```{r}
PAGE<-25
f_pona<-data.frame()

for (i in 1:PAGE) {
  url <- paste0("https://www.goodreads.com/shelf/show/politics?page=",i)
  print(url)
  page <- jump_to(pgsession, url)
  nds <- html_nodes(page, xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "bookTitle", " " ))]')
  politics_name <- html_text(nds)
  politics_name <- as.data.frame(politics_name)
  f_pona <- rbind(f_pona,politics_name)
}
```
put name and year in the same dataframe-politics
```{r}
f_politics <- cbind(f_pona, f_pubyr2)
```
find intersections
```{r}
#trim the line breaker
f_politics$politics_name<-trimws(f_politics$politics_name,which="both")
#delet whatever is in parentheses
f_politics$politics_name<-gsub("\\s*\\([^\\)]+\\)","",f_politics$politics_name)
#loop to get the number of race books among the top 200 for each year
popular_inter <- data.frame()
for (i in 1960:2019) {
  year_group <- which(f_politics$year==i)
  intersection <- intersect(f_top$top_name,f_politics$politics_name[year_group])
  intersection <- as.data.frame(length(intersection))
  popular_inter <- rbind(popular_inter,intersection)
}
names(popular_inter)<-"number"
#calculate the proportion
popular_inter$proportion <- popular_inter$number/200
#append column of year
popular_inter$year <- c(1960:2019)
popular_inter$year <- as.factor(popular_inter$year)
#plot of the proportion
ggplot(popular_inter,aes(x=year,y=proportion,fill=proportion))+
  geom_col(width=1)+
  scale_fill_continuous(low="light blue", high="blue")+
  scale_x_discrete(breaks = seq(1960, 2020, by = 5)) +
  ylim(0,0.1)+
  theme(axis.text.x = element_text(angle = 45))
```
intersection: politics and race
```{r}
pora_inter <- data.frame()
for (i in 1960:2019) {
  year_group_ra <- which(f_race$year==i)
  year_group_po <- which(f_politics$year==i)
  intersection <- intersect(f_race$race_name[year_group_ra],f_politics$politics_name[year_group_po])
  intersection <- as.data.frame(length(intersection))
  pora_inter <- rbind(pora_inter,intersection)
}
names(pora_inter)<-"number"
#calculate the proportion
pora_inter$proportion <- pora_inter$number/200
#append column of year
pora_inter$year <- c(1960:2019)
pora_inter$year <- as.factor(pora_inter$year)
#plot of the proportion
ggplot(pora_inter,aes(x=year,y=proportion))+
  geom_bar(stat="identity")
ggplot(pora_inter,aes(x=year,y=proportion,fill=proportion))+
  geom_col(width=1)+
  scale_fill_continuous(low="light blue", high="blue")+
  scale_x_discrete(breaks = seq(1960, 2020, by = 5)) +
  ylim(0,0.1)+
  theme(axis.text.x = element_text(angle = 45))
```
Final Graph: intersection of books about race and politics with top200 booklist in the same plot
```{r}
popular_inter1$group<-"race"
popular_inter$group<-"politics"
combine<-rbind(popular_inter1, popular_inter)

ggplot(combine,aes(x=year,y=proportion,group=group,color=group))+
  geom_line()+
  scale_x_discrete(breaks = seq(1960, 2020, by = 5)) +
  ylim(0,0.1)+
  theme(axis.text.x = element_text(angle = 45))
```

